---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, message = FALSE, eval = FALSE, echo = FALSE}
install.packages('ArchaeoPhases', repos = "http://cran.us.r-project.org")
```

```{r warning=FALSE, message=FALSE, echo = FALSE}
library(tidyverse)
library(lubridate)
library(rvest)
library(ggplot2)
library(stats)
library(ArchaeoPhases)
```

# Appendix

Load the FiveThirtyEight president polls from [https://projects.fivethirtyeight.com/polls/data/president_polls.csv](https://projects.fivethirtyeight.com/polls/data/president_polls.csv) into an object called `president_polls`. Remove polls that are not part of the 2024 president elections. Keep only the polls that are "general" elections and target a specific state. Also, only keep polls that have Biden or Trump as the candidate.

```{r}
president_polls <- read_csv("https://projects.fivethirtyeight.com/polls/data/president_polls.csv")
president_polls <- president_polls |> filter(cycle == 2024 & stage == "general" & !is.na(state) & candidate_name %in% c("Joe Biden", "Donald Trump"))
president_polls
```

Look at a table of the `fte_grade` given to each of the polls. Filter the `president_polls` data set to keep only polls with a grade in the top 10 grades, up to C.

```{r}
president_polls <- president_polls |> filter(fte_grade <= "C" | fte_grade == "C+")
president_polls
```

Make a table showing how many states have available poll data and what race_id is it.

```{r}
temp <- president_polls |> distinct(across(c(race_id, state, candidate_id)))
temp <- temp |> group_by(race_id, state) |> summarize(candidates = length(candidate_id))
temp
```

To make the following processing easier, change the name of the candidates to "Biden" and “Trump" respectively.

```{r}
president_polls["candidate_name"][president_polls["candidate_name"] == "Joe Biden"] = "Biden"
president_polls["candidate_name"][president_polls["candidate_name"] == "Donald Trump"] = "Trump"
```

Identify the columns we need to remove for `pivot_wider` to work and remove them. After removing these columns, use `pivot_wider` to create columns called `Biden` and `Trump` for the Democrat and Republican poll percentage, respectively.

```{r}
president_polls <- president_polls |> select(c(-answer, -candidate_id, -party)) |> pivot_wider(names_from = "candidate_name", values_from = "pct")
president_polls
```

Notice that there are some rows that either miss record for the supporting rate for Biden or the supporting rate for Trump. Let's show these rows.

```{r}
temp <- president_polls |> filter(is.na(Biden) | is.na(Trump)) |> select(pollster_id, question_id, Biden, Trump)
temp
```

According to the above table, we can see that it is the different question_id that prevent the rows from merging. Since there are not too many rows affected by this and there are only the supporting rate for Biden for most rows, I decide to remove all rows with missing supporting rate for either candidate.

```{r}
president_polls <- president_polls |> filter(!is.na(Biden) & !is.na(Trump))
```

Now define a new column `spread` as `(Biden-Trump)/100`.

```{r}
president_polls <- president_polls |> mutate(spread = (Biden - Trump) / 100)
president_polls
```

Use the `mdy` function in the `lubridate` package to mutate the start and end date columns in place to be Dates instead of characters.

```{r}
president_polls["start_date"] <- mdy(president_polls$start_date)
president_polls["end_date"] <- mdy(president_polls$end_date)
president_polls
```

Now let's see how many polls we have for each state.

```{r}
president_polls |> group_by(state) |> summarize(count = length(state)) |> arrange(-count)
```

For states with available polls > 6, let's see how Biden and Trump's supporting rates change with respect to time.

```{r}
get_plot <- function(target_state){
  president_polls |> filter(state == target_state) |> ggplot(aes(x = start_date)) + geom_line(aes(y = Biden, colour = "Biden")) + geom_line(aes(y = Trump, colour = "Trump")) + xlab("Date") + ylab("Supporting rate") + ggtitle(target_state) + scale_colour_manual("", 
                      breaks = c("Biden", "Trump"),
                      values = c("blue", "red")) + theme(legend.key.size = unit(0.5, 'cm'), #change legend key size
        legend.key.height = unit(0.5, 'cm'), #change legend key height
        legend.key.width = unit(0.5, 'cm'), #change legend key width
        legend.title = element_text(size=20), #change legend title font size
        legend.text = element_text(size=10), #change legend text font size
        plot.title = element_text(hjust = 0.5)) 
}
```
```{r}
get_plot("Georgia")
```
```{r}
get_plot("Missouri")
```
```{r}
get_plot("Arizona")
```
```{r}
get_plot("New Hampshire")
```
```{r}
get_plot("Pennsylvania")
```
```{r}
get_plot("Ohio")
```

**According to the plots above, we can see that Trump is leading Biden in all states except New Hampshire when it come to the midterm election**

Now calculate the mean and the standard deviation of the spread for each state. Since there are still two years before the 2024 presidential election and I do not have a lot of polls for some states, I decide to keep all the states that I have at least two polls available.

```{r}
results <- president_polls |> group_by(race_id, state) |> summarize(average = mean(spread), standard_deviation = sd(spread), count = length(unique(poll_id))) |> select(race_id, state, average, standard_deviation, count) |> filter(count > 1)
results 
```

Plot the results with confidence intervals assuming the average is t-distributed. Put state on the y-axis and the average along the x-axis, using `geom_errorbar` to specify the confidence interval width about the average. Order the states by the absolute value of the average difference so the closer races are on top. Color them blue if Biden is winning and red if Trump is winning (don't show a legend).

```{r}
# Calculate 95% confidence interval
standard_error <- results$standard_deviation / sqrt(results$count)
alpha <- 0.05
degrees_of_freedom <- results$count - 1
t_score = qt(p=alpha/2, df=degrees_of_freedom,lower.tail=F)
margin_error <- t_score * standard_error
lower_bound <- results$average - margin_error
upper_bound <- results$average + margin_error
results |> ggplot(aes(x = average, y = reorder(state, -abs(average)))) + geom_point(aes(col = ifelse(average > 0, "blue", ifelse(average < 0, "red", "black"))), size = 3) + geom_errorbar(aes(xmin = lower_bound, xmax = upper_bound, colour = ifelse(average > 0, "blue", ifelse(average < 0, "red", "black")))) + scale_color_manual(values = c("blue" = "blue", "red"="red", "black"="black")) + xlab("Average") + ylab("States") + ggtitle("Poll by state") + theme(legend.position="none") + theme(plot.title = element_text(hjust = 0.5))
```

Now comes the important part. We need to predict how many electoral votes that Biden will win in the 2024 presidential election, so we need a table showing how many electoral votes that each state has. According to the demographic census in 2020, the electoral votes for each state have been updated and will firstly take effect in the 2024 presidential election. Below is the how I extract the table from the website.

The table is located in the following website: [https://state.1keydata.com/state-electoral-votes.php](https://state.1keydata.com/state-electoral-votes.php)

```{r}
url <- "https://state.1keydata.com/state-electoral-votes.php"
dat <- read_html(url)
```

Use the `html_nodes` function and the `table` node to extract the table we need.

```{r}
nodes <- dat |> html_nodes("table")
nodes
```

Our table of interest is the fourth table, so we extract it below and do some modifications.

```{r}
votes <- html_table(nodes[[4]], header = TRUE) |> select(2, 3)
colnames(votes)[1] <- 'state'
colnames(votes)[2] <- 'electoral_votes'
votes
```

The table above does not include District of Columbia, which has 3 electoral votes, so I add it into the table.

```{r}
add <- data.frame("District of Columbia", 3)
names(add)=c("state","electoral_votes")
votes <- votes |> rbind(add) |> arrange(-electoral_votes, state)
votes
```

Let's see how many electoral votes will be needed to win the 2024 presidential election.

```{r}
total <- sum(votes$electoral_votes)
print(paste(total %/% 2 + 1, " electoral votes are needed to win the 2024 presidential election"))
```

Join this table and the "results" table above.

```{r}
results <- left_join(results, votes, by = "state")
results
```

Now implement a Bayesian approach. Firstly, let's use a basic prior ($\mu = 0$, $\sigma = 0.04$) assuming no knowledge about the historical elections and the election trends.

```{r}
mu <- 0.00
tau <- 0.04
results_no_bias <- results |> mutate(prior_average = mu, 
                             prior_sd = tau, 
                             sigma = standard_deviation/sqrt(count),
                             B = sigma^2 / (sigma^2 + tau^2),
                             posterior_mean = B * mu + (1 - B) * average,
                             posterior_sd = sqrt(1/ (1/sigma^2 + 1/tau^2)),
                             Biden_win_prob = 1-pnorm(0, posterior_mean, posterior_sd)) |> select(-sigma, -B)
results_no_bias
```

Show a table of states with the probability (in percentages) of the Biden winning ordered by that probability. If a probability is larger than 99%, show >99%, and if smaller than 1% then show <1%.

```{r}
temp <- results_no_bias |> select(state, Biden_win_prob) |> mutate(Biden_win = paste(as.character(round(100 * Biden_win_prob, 2)), "%")) |> arrange(-Biden_win_prob)
temp["Biden_win"][temp["Biden_win_prob"] > 0.99] <- "> 99 %"
temp["Biden_win"][temp["Biden_win_prob"] < 0.01] <- "< 1 %"
temp |> select(state, Biden_win)
```

Note that there are several states that do not appear in the "results_no_bias" table above. This is because the election in these states are usually certain. As a result, I use the following strategy. If the state is not in the "results_no_bias" table, I consider the election results in these states are certain and count the electoral votes towards their party affiliation. To get the party affiliation, I extracted the following table from [https://www.pewresearch.org/religion/religious-landscape-study/compare/party-affiliation/by/state/](https://www.pewresearch.org/religion/religious-landscape-study/compare/party-affiliation/by/state/) using the similar web scraping strategy mentioned above.

```{r}
url <- "https://www.pewresearch.org/religion/religious-landscape-study/compare/party-affiliation/by/state/"
dat <- read_html(url)
nodes <- dat |> html_nodes("table")
nodes
```

```{r}
affiliation <- html_table(nodes[[1]], header = TRUE)
colnames(affiliation)[1] <- 'state'
colnames(affiliation)[2] <- 'Rep'
colnames(affiliation)[3] <- 'None'
colnames(affiliation)[4] <- 'Dem'
affiliation
```

The table above shows the percentage of the party affiliation that the adults in the states consider themselves as. Here, I consider the state as a "Blue state" if it has more adults that think they are Democrats. and a "Red state" if it has more adults that think they are Republicans. Otherwise, I label the state as a "White state" since it has no affiliation. As a result, I get the following table.

```{r}
affiliation <- affiliation |> mutate(affiliation = ifelse(Dem > Rep, "B", ifelse(Rep > Dem, "R", "W")))  |> select(-"Sample	size")
affiliation
```

Now let's only keep the states that are not in the "results_no_bias" table since we will use a Monte Carlo simulation to determine the election results in those states.

```{r}
affiliation <- affiliation |> filter(!state %in% results_no_bias$state)
affiliation
```

Let's assume that Biden manage to win in all blue states that remain in the "affiliation" table and lose in all red states that remain in the "affiliation" table. Before using the Monte Carlo simulation, let's see how many electoral votes that Biden has already won.

```{r}
affiliation <- left_join(affiliation, votes, by = "state")
total_votes <- affiliation |> filter(affiliation == "B") |> pull(electoral_votes)
Biden_win_electoral_votes <- sum(total_votes)
print(paste("Based on the assumption, Biden has already won ", Biden_win_electoral_votes, " electoral votes"))
```

Now Create a Monte Carlo simulation of each state's election using number of simulations `B=10000`, where each simulation outputs the total number of electoral votes that Biden will win in the 2024 presidential election. Assume each state's election is normally distributed with the posterior mean and standard error in the `results_no_bias` data frame. Show a histogram of the results for the number of electoral votes.

```{r}
# We now start the Monte Carlo simulation based on the above result
set.seed(1)
B <- 10000
simulation <- function(){
  temp <- c()
  for (i in 1:length(results_no_bias$state)){
    temp <- c(sample(c(1, 0), size = 1, replace = TRUE, prob = c(pnorm(0, mean = results_no_bias$posterior_mean[i], sd = results_no_bias$posterior_sd[i], lower.tail = FALSE), pnorm(0, mean = results_no_bias$posterior_mean[i], sd = results_no_bias$posterior_sd[i], lower.tail = TRUE))) * results_no_bias$electoral_votes[i], temp)
  }
  sum(temp) + Biden_win_electoral_votes
}
result_mc <- data.frame(electoral_votes_count = replicate(B, simulation()))
result_mc |> ggplot(aes(x = electoral_votes_count)) + geom_histogram() + xlab("Number of electoral votes Biden wins") + ylab("Count") + ggtitle("Presidential Election Possibilities") + theme(plot.title = element_text(hjust = 0.5))
```

Show the probability that Biden wins the 2024 presidential election, assuming win means 270 or more electoral votes.

```{r}
print(paste("According to the above plot, the probability that Biden wins the 2024 presidential election is ", round(mean(result_mc >= 270) * 100, 2), "%"))
```

Give the 80% credible interval of the 2024 presidential election.

```{r}
# I use the credible_interval function in the ArchaeoPhases package to get the 80% credible interval
result <- credible_interval(data = result_mc |> pull(electoral_votes_count), level = 0.8, round_to = 0)

print(paste("According to the redible_interval function in the ArchaeoPhases package, the 80% credible interval should be [", result$ci[1], ", ", result$ci[2], "]"))
```

Now including a bias with average `bias_avg` and standard error of  `bias_sd` to add the uncertainty to the election.

```{r}
# According to section 16.8.4 in the book, I set bias_sd to be 0.03. I also set bias_avg to be 0.00 since I do not think the bias will go towards either party
mu <- 0.00
tau <- 0.04
bias_avg <- 0.00
bias_sd <- 0.03
results_bias <- results |> mutate(prior_average = mu, 
                             prior_sd = tau, 
                             sigma = sqrt(standard_deviation^2/count  + bias_sd^2),
                             B = sigma^2 / (sigma^2 + tau^2),
                             posterior_mean = B * mu + (1 - B) * average,
                             posterior_sd = sqrt(1/ (1/sigma^2 + 1/tau^2)),
                             Biden_win_prob = 1-pnorm(0, posterior_mean, posterior_sd)) |> select(-sigma, -B)
results_bias
```

Print out the table of states with Biden's probabilities of winning. If a probability is larger than 99%, show >99%, and if smaller than 1% then show <1%.

```{r}
# Print out the table of states with Biden's probabilities of winning
temp <- results_bias |> select(state, Biden_win_prob) |> mutate(Biden_win = paste(as.character(round(100 * Biden_win_prob, 2)), "%"))
temp["Biden_win"][temp["Biden_win_prob"] > 0.99] <- "> 99 %"
temp["Biden_win"][temp["Biden_win_prob"] < 0.01] <- "< 1 %"
temp |> select(state, Biden_win)
```

Now run a simulation to get a final answer for the total of electoral votes that Biden will win in the 2024 presidential election.

```{r}
# Use the updated results table to perform the Monte Carlo simulation
set.seed(1)
B <- 10000
simulation <- function(){
  temp <- c()
  for (i in 1:length(results_bias$state)){
    temp <- c(sample(c(1, 0), size = 1, replace = TRUE, prob = c(pnorm(0, mean = results_bias$posterior_mean[i], sd = results_bias$posterior_sd[i], lower.tail = FALSE), pnorm(0, mean = results_bias$posterior_mean[i], sd = results_bias$posterior_sd[i], lower.tail = TRUE))) * results_bias$electoral_votes[i], temp)
  }
  sum(temp) + Biden_win_electoral_votes
}
result_mc <- data.frame(electoral_votes_count = replicate(B, simulation()))
result_mc |> ggplot(aes(x = electoral_votes_count)) + geom_histogram() + xlab("Number of electoral votes Biden wins") + ylab("Count") + ggtitle("Presidential Election Possibilities") + theme(plot.title = element_text(hjust = 0.5))
```

Let's see now how likely Biden will win against Trump with the new posterior.

```{r}
print(paste("According to the above plot, the probability that the Democrats control the Senate is ", round(mean(result_mc >= 270) * 100, 2), "%"))
```

Give the 80% credible interval of the 2024 presidential election.

```{r}
# I use the credible_interval function in the ArchaeoPhases package to get the 80% credible interval
result <- credible_interval(data = result_mc |> pull(electoral_votes_count), level = 0.8, round_to = 0)

print(paste("According to the redible_interval function in the ArchaeoPhases package, the 80% credible interval should be [", result$ci[1], ", ", result$ci[2], "]"))
```

According to the above results, by adding a bias term, the election results become more uncertain and Biden's winning rate increases. However, Biden will still suffer great loss regardless of the bias term is added or not. However, the above results are based on the prior that has no knowledge about the previous presidential elections and we are subjected to limited poll data.

The previous analysis may give a more plausible prediction if we can set a better prior for each state. To achieve this, we use the poll results from the 2020 presidential election. This is because this election is also between Biden and Trump. First, let's get the data from FiveThirtyEight again and conduct necessary data wrangling. The data is available at [https://projects.fivethirtyeight.com/polls/data/president_polls_historical.csv](https://projects.fivethirtyeight.com/polls/data/president_polls_historical.csv)

```{r}
president_polls_2020 <- read_csv("https://projects.fivethirtyeight.com/polls/data/president_polls_historical.csv")
president_polls_2020 <- president_polls_2020 |> filter(cycle == 2020 & stage == "general" & !is.na(state) & candidate_name %in% c("Joe Biden", "Donald Trump") & fte_grade <= "C" | fte_grade == "C+")
president_polls_2020["start_date"] <- mdy(president_polls_2020$start_date)
president_polls_2020["end_date"] <- mdy(president_polls_2020$end_date)
president_polls_2020["election_date"] <- mdy(president_polls_2020$election_date)
president_polls_2020["candidate_name"][president_polls_2020["candidate_name"] == "Joe Biden"] = "Biden"
president_polls_2020["candidate_name"][president_polls_2020["candidate_name"] == "Donald Trump"] = "Trump"
president_polls_2020
```

Use "pivot_wider" to merge the rows and create the "spread" column that indicates the supporting rate spread between Biden and Trump.

```{r}
president_polls_2020 <- president_polls_2020 |> select(c(-answer, -candidate_id, -party)) |> pivot_wider(names_from = "candidate_name", values_from = "pct") |> mutate(spread = (Biden - Trump) / 100) |> filter(!is.na(spread))
president_polls_2020
```

After consideration, I think a good prior should be set as a normal distribution since the election result is usually normally distributed. The mean of prior should be the average of the spread within the final month of the election since we want to predict the results that are close to the election day. The standard deviation of the prior should be variance between 2018 to 2020. This is because the earliest poll data from FiveThirtyEight is since 2018 and I want the standard deviation to be relatively large to account for uncertainties since I am predicting the 2024 presidential election using the data in 2022 and a lot can happen during the next two years. Let's find the mean and the standard deviation of the prior following the definitions above.

```{r}
temp_1 <- president_polls_2020 |> filter(start_date > election_date - months(1)) |> group_by(state) |> summarize(prior_average = mean(spread)) |> filter(!is.na(prior_average))
temp_1
```

```{r}
temp_2 <- president_polls_2020 |> group_by(state) |> summarize(prior_sd = sd(spread)) |> filter(!is.na(prior_sd))
temp_2
```

Note that there are more than 50 rows in the tables above. This is because there are polls for some regions of one state, but this does not matter since we only want to have the prior mean and standard deviation that are in "results" table above. Now we add what we have found into the "results" table by joining it with the "temp_1" and "temp_2" table.

```{r}
updated_results <- left_join(results, temp_1, by = "state")
updated_results <- left_join(updated_results, temp_2, by = "state")
updated_results
```

Perform a Bayesian approach using the updated prior average and standard deviation. First, let's assume no bias term.

```{r}
updated_results_no_bias <- updated_results |> mutate(sigma = standard_deviation/sqrt(count),
                             B = sigma^2 / (sigma^2 + prior_sd^2),
                             posterior_mean = B * prior_average + (1 - B) * average,
                             posterior_sd = sqrt(1/ (1/sigma^2 + 1/prior_sd^2)),
                             Biden_win_prob = 1-pnorm(0, posterior_mean, posterior_sd)) |> select(-sigma, -B)
updated_results_no_bias
```

Print out the table of states with Biden's probabilities of winning. If a probability is larger than 99%, show >99%, and if smaller than 1% then show <1%.

```{r}
temp <- updated_results_no_bias |> select(state, Biden_win_prob) |> mutate(Biden_win = paste(as.character(round(100 * Biden_win_prob, 2)), "%")) |> arrange(-Biden_win_prob)
temp["Biden_win"][temp["Biden_win_prob"] > 0.99] <- "> 99 %"
temp["Biden_win"][temp["Biden_win_prob"] < 0.01] <- "< 1 %"
temp |> select(state, Biden_win)
```

Create a Monte Carlo simulation using the updated posterior average and standard deviation.

```{r}
# We now start the Monte Carlo simulation based on the above result
set.seed(1)
B <- 10000
simulation <- function(){
  temp <- c()
  for (i in 1:length(updated_results_no_bias$state)){
    temp <- c(sample(c(1, 0), size = 1, replace = TRUE, prob = c(pnorm(0, mean = updated_results_no_bias$posterior_mean[i], sd = updated_results_no_bias$posterior_sd[i], lower.tail = FALSE), pnorm(0, mean = updated_results_no_bias$posterior_mean[i], sd = updated_results_no_bias$posterior_sd[i], lower.tail = TRUE))) * updated_results_no_bias$electoral_votes[i], temp)
  }
  sum(temp) + Biden_win_electoral_votes
}
result_mc <- data.frame(electoral_votes_count = replicate(B, simulation()))
result_mc |> ggplot(aes(x = electoral_votes_count)) + geom_histogram() + xlab("Number of electoral votes Biden wins") + ylab("Count") + ggtitle("Presidential Election Possibilities") + theme(plot.title = element_text(hjust = 0.5))
```

Let's see now how likely Biden will win against Trump with the new posterior.

```{r}
print(paste("According to the above plot, the probability that Biden wins the 2024 presidential election is ", round(mean(result_mc >= 270) * 100, 2), "%"))
```

Give the 80% credible interval of the 2024 presidential election.

```{r}
# I use the credible_interval function in the ArchaeoPhases package to get the 80% credible interval
result <- credible_interval(data = result_mc |> pull(electoral_votes_count), level = 0.8, round_to = 0)

print(paste("According to the redible_interval function in the ArchaeoPhases package, the 80% credible interval should be [", result$ci[1], ", ", result$ci[2], "]"))
```

Now including a bias with average `bias_avg` and standard error of  `bias_sd` to add the uncertainty to the election. Print out the table of states with Biden's probabilities of winning.

```{r}
# According to section 16.8.4 in the book, I set bias_sd to be 0.03. I also set bias_avg to be 0.00 since I do not think the bias will go towards either party
bias_avg <- 0.00
bias_sd <- 0.03
updated_results_bias <- updated_results |> mutate(sigma = sqrt(standard_deviation^2/count  + bias_sd^2),
                             B = sigma^2 / (sigma^2 + prior_sd^2),
                             posterior_mean = B * prior_average + (1 - B) * average,
                             posterior_sd = sqrt(1/ (1/sigma^2 + 1/prior_sd^2)),
                             Biden_win_prob = 1-pnorm(0, posterior_mean, posterior_sd)) |> select(-sigma, -B)


updated_results_bias
```

Print out the table of states with Biden's probabilities of winning. If a probability is larger than 99%, show >99%, and if smaller than 1% then show <1%.

```{r}
# Print out the table of states with Biden's probabilities of winning
temp <- updated_results_bias |> select(state, Biden_win_prob) |> mutate(Biden_win = paste(as.character(round(100 * Biden_win_prob, 2)), "%"))
temp["Biden_win"][temp["Biden_win_prob"] > 0.99] <- "> 99 %"
temp["Biden_win"][temp["Biden_win_prob"] < 0.01] <- "< 1 %"
temp |> select(state, Biden_win)
```

Create a Monte Carlo simulation using the updated posterior average and standard deviation.

```{r}
# We now start the Monte Carlo simulation based on the above result
set.seed(1)
B <- 10000
simulation <- function(){
  temp <- c()
  for (i in 1:length(updated_results_bias$state)){
    temp <- c(sample(c(1, 0), size = 1, replace = TRUE, prob = c(pnorm(0, mean = updated_results_bias$posterior_mean[i], sd = updated_results_bias$posterior_sd[i], lower.tail = FALSE), pnorm(0, mean = updated_results_bias$posterior_mean[i], sd = updated_results_bias$posterior_sd[i], lower.tail = TRUE))) * updated_results_bias$electoral_votes[i], temp)
  }
  sum(temp) + Biden_win_electoral_votes
}
result_mc <- data.frame(electoral_votes_count = replicate(B, simulation()))
result_mc |> ggplot(aes(x = electoral_votes_count)) + geom_histogram() + xlab("Number of electoral votes Biden wins") + ylab("Count") + ggtitle("Presidential Election Possibilities") + theme(plot.title = element_text(hjust = 0.5))
```

Let's see now how likely Biden will win against Trump with the new posterior.

```{r}
print(paste("According to the above plot, the probability that Biden wins the 2024 presidential election is ", round(mean(result_mc >= 270) * 100, 2), "%"))
```

Give the 80% credible interval of the 2024 presidential election.

```{r}
# I use the credible_interval function in the ArchaeoPhases package to get the 80% credible interval
result <- credible_interval(data = result_mc |> pull(electoral_votes_count), level = 0.8, round_to = 0)

print(paste("According to the redible_interval function in the ArchaeoPhases package, the 80% credible interval should be [", result$ci[1], ", ", result$ci[2], "]"))
```

According to the results above, we can see that Biden's winning rate increases greatly against Trump in both cases (with or without the bias), and it suggests that it is Biden who are supposed to win the election, which overturns the results generated using our previous prior. This is anticipated since for the last month of the election, people have seen Trump's performance for the four years but they have not seen Biden's performance. As a result, the polls will go against Trump and favor Biden. However, in the 2024 presidential election, people will see how Biden performed in the four years and the polls are more likely to go against Biden.

Let's have another try by assuming that in 2024, people have seen both Biden and Trump's performance during their presidency and their favor towards both candidates is mitigated, namely they do not favor strongly for either Biden or Trump. Based on this assumption, we can have another prior with the prior mean to be the average of the current poll average and the average of the polls in the last month of the 2020 presidential election, with the prior standard deviation the same as the one in our second try.

```{r}
updated_results["prior_average"] <- (updated_results$average + updated_results$prior_average) / 2
updated_results
```

Perform a Bayesian approach using the updated prior average and standard deviation. First, let's assume no bias term.

```{r}
updated_results_no_bias <- updated_results |> mutate(sigma = standard_deviation/sqrt(count),
                             B = sigma^2 / (sigma^2 + prior_sd^2),
                             posterior_mean = B * prior_average + (1 - B) * average,
                             posterior_sd = sqrt(1/ (1/sigma^2 + 1/prior_sd^2)),
                             Biden_win_prob = 1-pnorm(0, posterior_mean, posterior_sd)) |> select(-sigma, -B)
updated_results_no_bias
```

Print out the table of states with Biden's probabilities of winning. If a probability is larger than 99%, show >99%, and if smaller than 1% then show <1%.

```{r}
temp <- updated_results_no_bias |> select(state, Biden_win_prob) |> mutate(Biden_win = paste(as.character(round(100 * Biden_win_prob, 2)), "%")) |> arrange(-Biden_win_prob)
temp["Biden_win"][temp["Biden_win_prob"] > 0.99] <- "> 99 %"
temp["Biden_win"][temp["Biden_win_prob"] < 0.01] <- "< 1 %"
temp |> select(state, Biden_win)
```

Create a Monte Carlo simulation using the updated posterior average and standard deviation.

```{r}
# We now start the Monte Carlo simulation based on the above result
set.seed(1)
B <- 10000
simulation <- function(){
  temp <- c()
  for (i in 1:length(updated_results_no_bias$state)){
    temp <- c(sample(c(1, 0), size = 1, replace = TRUE, prob = c(pnorm(0, mean = updated_results_no_bias$posterior_mean[i], sd = updated_results_no_bias$posterior_sd[i], lower.tail = FALSE), pnorm(0, mean = updated_results_no_bias$posterior_mean[i], sd = updated_results_no_bias$posterior_sd[i], lower.tail = TRUE))) * updated_results_no_bias$electoral_votes[i], temp)
  }
  sum(temp) + Biden_win_electoral_votes
}
result_mc <- data.frame(electoral_votes_count = replicate(B, simulation()))
result_mc |> ggplot(aes(x = electoral_votes_count)) + geom_histogram() + xlab("Number of electoral votes Biden wins") + ylab("Count") + ggtitle("Presidential Election Possibilities") + theme(plot.title = element_text(hjust = 0.5))
```

Let's see now how likely Biden will win against Trump with the new posterior.

```{r}
print(paste("According to the above plot, the probability that Biden wins the 2024 presidential election is ", round(mean(result_mc >= 270) * 100, 2), "%"))
```

Give the 80% credible interval of the 2024 presidential election.

```{r}
# I use the credible_interval function in the ArchaeoPhases package to get the 80% credible interval
result <- credible_interval(data = result_mc |> pull(electoral_votes_count), level = 0.8, round_to = 0)

print(paste("According to the redible_interval function in the ArchaeoPhases package, the 80% credible interval should be [", result$ci[1], ", ", result$ci[2], "]"))
```

Now including a bias with average `bias_avg` and standard error of  `bias_sd` to add the uncertainty to the election. Print out the table of states with Biden's probabilities of winning.

```{r}
# According to section 16.8.4 in the book, I set bias_sd to be 0.03. I also set bias_avg to be 0.00 since I do not think the bias will go towards either party
bias_avg <- 0.00
bias_sd <- 0.03
updated_results_bias <- updated_results |> mutate(sigma = sqrt(standard_deviation^2/count  + bias_sd^2),
                             B = sigma^2 / (sigma^2 + prior_sd^2),
                             posterior_mean = B * prior_average + (1 - B) * average,
                             posterior_sd = sqrt(1/ (1/sigma^2 + 1/prior_sd^2)),
                             Biden_win_prob = 1-pnorm(0, posterior_mean, posterior_sd)) |> select(-sigma, -B)


updated_results_bias
```

Print out the table of states with Biden's probabilities of winning. If a probability is larger than 99%, show >99%, and if smaller than 1% then show <1%.

```{r}
# Print out the table of states with Biden's probabilities of winning
temp <- updated_results_bias |> select(state, Biden_win_prob) |> mutate(Biden_win = paste(as.character(round(100 * Biden_win_prob, 2)), "%"))
temp["Biden_win"][temp["Biden_win_prob"] > 0.99] <- "> 99 %"
temp["Biden_win"][temp["Biden_win_prob"] < 0.01] <- "< 1 %"
temp |> select(state, Biden_win)
```

Create a Monte Carlo simulation using the updated posterior average and standard deviation.

```{r}
# We now start the Monte Carlo simulation based on the above result
set.seed(1)
B <- 10000
simulation <- function(){
  temp <- c()
  for (i in 1:length(updated_results_bias$state)){
    temp <- c(sample(c(1, 0), size = 1, replace = TRUE, prob = c(pnorm(0, mean = updated_results_bias$posterior_mean[i], sd = updated_results_bias$posterior_sd[i], lower.tail = FALSE), pnorm(0, mean = updated_results_bias$posterior_mean[i], sd = updated_results_bias$posterior_sd[i], lower.tail = TRUE))) * updated_results_bias$electoral_votes[i], temp)
  }
  sum(temp) + Biden_win_electoral_votes
}
result_mc <- data.frame(electoral_votes_count = replicate(B, simulation()))
result_mc |> ggplot(aes(x = electoral_votes_count)) + geom_histogram() + xlab("Number of electoral votes Biden wins") + ylab("Count") + ggtitle("Presidential Election Possibilities") + theme(plot.title = element_text(hjust = 0.5))
```

Let's see now how likely Biden will win against Trump with the new posterior.

```{r}
print(paste("According to the above plot, the probability that Biden wins the 2024 presidential election is ", round(mean(result_mc >= 270) * 100, 2), "%"))
```

Give the 80% credible interval of the 2024 presidential election.

```{r}
# I use the credible_interval function in the ArchaeoPhases package to get the 80% credible interval
result <- credible_interval(data = result_mc |> pull(electoral_votes_count), level = 0.8, round_to = 0)

print(paste("According to the redible_interval function in the ArchaeoPhases package, the 80% credible interval should be [", result$ci[1], ", ", result$ci[2], "]"))
```

I believe that the result above is more reasonable. It is sometimes the case that a president lose the midterm election but win in the presidential election two years later. Since I think the polls in 2020 favor Biden and the polls in 2022 favor Trump, by mitigating the polls results, we should approach the results better.
